---
title: "Transfer analyses"
output: html_document
---

```{r}
library(tidyverse)
library(lme4)
library(lmerTest)
```

# load data
```{r}
num_runs = 200
num_layerss = 2:4
```

```{r message=F, warning=F}
d = data.frame()
for (run_i in 0:(num_runs-1)) {
  for (num_layers in num_layerss) {
    for (analogous in 0:1) {
      for (nonlinear in 0:1) {
        filename = sprintf("results/nlayer_%i_nonlinear_%i_analogous_%i_rseed_%i_loss_track.csv",
                           num_layers, nonlinear, analogous, run_i)
        this_d = read_csv(filename) %>%
          mutate(run=run_i, nonlinear=nonlinear==1, analogous=factor(analogous, levels=0:1, labels=c("non-analogous", "analogous")), num_layers=num_layers)
        d = bind_rows(d, this_d)
      }
    }
  }
}
```

```{r}
d = d %>%
  filter((num_layers==5 & epoch %% 500 == 0) | (num_layers==4 & epoch %% 200 == 0) | (num_layers == 3 & epoch %% 50 == 0) | num_layers == 2) %>%
  complete(run, analogous, nesting(nonlinear, num_layers, epoch), fill=list(MSE=0, d1_MSE=0)) 
```


# basic analysis
```{r}
theme_set(theme_bw() +
          theme(panel.grid=element_blank()))
```

```{r}
ggplot(data=d, aes(x=epoch, y=d1_MSE, color=analogous)) +
  geom_line(stat="summary", fun.y=median, size=1) +
  geom_line(aes(group=interaction(run, analogous)), alpha=0.1) +
  facet_wrap(.~nonlinear + num_layers, scales="free_x") +
  labs(y="Loss (L2)") +
  scale_color_brewer(palette="Set1")
```

```{r}
ggsave("plots/basic_transfer_comparison.png", width=8, height=4)
```

```{r}
learned_d = d %>%
  filter(d1_MSE < 0.05) %>%
  group_by(nonlinear, analogous, num_layers, run) %>%
  summarize(learned_epoch = min(epoch)) %>%
  ungroup()
```


```{r}
model = lmer(learned_epoch ~ analogous  * nonlinear + (1 | run), learned_d %>% filter(num_layers==3))

summary(model)
```

```{r}
learned_diff_d = learned_d %>%
  spread(analogous, learned_epoch) %>%
  mutate(learn_diff = `non-analogous` - `analogous`)
```

```{r}
ggplot(learned_d %>% filter(nonlinear, num_layers==3), aes(x=learned_epoch, color=analogous)) +
  geom_line(stat="density") +
  facet_grid(. ~ num_layers, scales="free")
```
```{r}
ggsave("plots/basic_transfer_stopping_times.png", width=8, height=4)
```


```{r}
ggplot(learned_diff_d %>% filter(nonlinear, num_layers==3), aes(x=learn_diff)) +
  geom_line(stat="density") +
  facet_grid(. ~ num_layers, scales="free") +
  labs(x="Amount longer it took to learn non-analogous")
```

```{r}
ggsave("plots/basic_transfer_stopping_time_differences.png", width=8, height=4)
```


```{r}
sorted_diff_d = arrange(learned_diff_d %>% filter(num_layers==3, nonlinear), desc(learn_diff))
head(sorted_diff_d)

tail(sorted_diff_d)

```

```{r}
ggplot(data=d %>% filter(run %in% c(80, 104, 107), num_layers==3, nonlinear), aes(x=epoch, y=d1_MSE, color=analogous)) +
  geom_line(aes(group=interaction(run, analogous))) +
  labs(y="Loss (L2)") +
  scale_color_brewer(palette="Set1")
```

Since we know in the 3 layer case that the transfer effect **must** be happening at the middle layer, since these are the only shared weights. This provides a nice target for analysis.

## why are some 4-layer linear runs showing effects in the other direction?


```{r}
ggplot(data=d %>% filter(run %in% c(98), num_layers==4, !nonlinear), aes(x=epoch, y=d1_MSE, color=analogous)) +
  geom_line(aes(group=interaction(run, analogous))) +
  labs(y="Loss (L2)") +
  scale_color_brewer(palette="Set1")
```

```{r}
ggplot(data=d %>% filter(run %in% c(98), num_layers==4, !nonlinear), aes(x=epoch, y=MSE, color=analogous)) +
  geom_line(aes(group=interaction(run, analogous))) +
  labs(y="Loss (L2)") +
  scale_color_brewer(palette="Set1")
```

Might just be single-domain weirdness and the exaggeration that happens with depth... Should investigate further.


# Investigating how the penultimate layer weights change

```{r}
detailed_runs = 0:135
num_layers = 3
```

```{r message=FALSE}
simil_d = data.frame()
sv_d = data.frame()
proj_d = data.frame()
for (run_i in detailed_runs) {
  for (analogous in 0:1) {
    for (nonlinear in c(1)) {
      filename = sprintf("results/nlayer_%i_nonlinear_%i_analogous_%i_rseed_%i_penultimate_simil_track.csv",
                         num_layers, nonlinear, analogous, run_i)
      if (!file.exists(filename)) {
        next
      }
      this_d = read_csv(filename) %>%
        mutate(run=run_i, nonlinear=nonlinear==1, analogous=factor(analogous, levels=0:1, labels=c("non-analogous", "analogous")), num_layers=num_layers)
      simil_d = bind_rows(simil_d, this_d)
      
      filename = sprintf("results/nlayer_%i_nonlinear_%i_analogous_%i_rseed_%i_penultimate_S_track.csv",
                         num_layers, nonlinear, analogous, run_i)
      this_d = read_csv(filename) %>%
        mutate(run=run_i, nonlinear=nonlinear==1, analogous=factor(analogous, levels=0:1, labels=c("non-analogous", "analogous")), num_layers=num_layers)
      sv_d = bind_rows(sv_d, this_d)

      filename = sprintf("results/nlayer_%i_nonlinear_%i_analogous_%i_rseed_%i_penultimate_proj_track.csv",
                         num_layers, nonlinear, analogous, run_i)
      this_d = read_csv(filename) %>%
        mutate(run=run_i, nonlinear=nonlinear==1, analogous=factor(analogous, levels=0:1, labels=c("non-analogous", "analogous")), num_layers=num_layers)
      proj_d = bind_rows(proj_d, this_d)
    }
  }
}
```


## projections
```{r}
part_summarized_proj_d = proj_d %>%
  mutate(mode_rank = mode_j,
         abs_projection=projection) %>%
  select(-projection) %>%
  spread(rep_i, abs_projection) %>%
  group_by(run, analogous, epoch, mode_rank) %>%
  summarize(abs_proj_1 = `0` + `1` + `2`,
            abs_proj_2 = `3` + `4` + `5`,
            abs_proj_diff = abs(abs_proj_1 - abs_proj_2)) %>% # how much is the mode biased towards one domain?
  ungroup() 

summarized_proj_d = part_summarized_proj_d %>%
  group_by(analogous, epoch, mode_rank) %>%
  summarize(mean_abs_proj_diff = mean(abs_proj_diff), sd_abs_proj_diff = sd(abs_proj_diff))

final_proj_d = part_summarized_proj_d %>%
  group_by(run, analogous, mode_rank) %>%
  filter(epoch == max(epoch)) %>%
  ungroup()
```

```{r}
ggplot(summarized_proj_d %>% filter(mode_rank < 2), aes(x=epoch, y=mean_abs_proj_diff, color=analogous)) +
  geom_line() +
  scale_color_brewer(palette="Set1") +
  facet_grid(~mode_rank)
```

```{r}
ggplot(final_proj_d, aes(x=abs_proj_diff, color=analogous)) +
  geom_density() +
  scale_color_brewer(palette="Set1") +
  facet_grid(~mode_rank)
```

## similarities 

```{r}
part_summarized_simil_d = simil_d %>%
  rename(cosine_distance=cosine_similarity) %>% # fix incorrect label in data output
  mutate(type= case_when(rep_j == rep_i + 3 ~ "between_analog", 
                         rep_i < 3 & rep_j >= 3 ~ "between_nonanalog",
                         rep_i < 3 & rep_j < 3 ~ "within_d1",
                         T ~ "within_d2"),
         abs_cosine=abs(1-cosine_distance)) %>%
  group_by(run, analogous, epoch, type) %>%
  summarize(mean_abs_cosine = mean(abs_cosine),
            mean_cosine_distance = mean(cosine_distance))

summarized_simil_d = part_summarized_simil_d %>%
  group_by(epoch, analogous, type) %>%
  summarize(mean_abs_cosine = mean(mean_abs_cosine),
            mean_cosine_distance = mean(mean_cosine_distance))

final_simil_d = part_summarized_simil_d %>%
  group_by(run, analogous, type) %>%
  filter(epoch == max(epoch)) %>%
  ungroup()
```

```{r}
ggplot(final_simil_d, aes(x=mean_cosine_distance, color=analogous)) +
  geom_density() +
  scale_color_brewer(palette="Set1") +
  facet_grid(~type)
```


```{r}
ggplot(final_simil_d, aes(x=mean_abs_cosine, color=analogous)) +
  geom_density() +
  scale_color_brewer(palette="Set1") +
  facet_grid(~type)
```

## singular values
```{r}
sv_d = sv_d %>%
  rename(mode_rank=rank) %>%
  group_by(run, mode_rank, analogous) %>%
  mutate(norm_S = S/max(S)) %>%
  ungroup()

summarized_sv_d = sv_d %>%
  group_by(analogous, epoch, mode_rank) %>%
  summarize(mean_norm_S=mean(norm_S), mean_S=mean(S))

final_sv_d = sv_d %>%
  group_by(analogous, run, mode_rank) %>% 
  filter(epoch == max(epoch)) %>%
  ungroup()

early_sv_d = sv_d %>%
  group_by(analogous, run, mode_rank) %>% 
  filter(epoch == 5000) %>%
  select(-epoch) %>%
  ungroup()
```

```{r}
ggplot(final_sv_d, aes(x=S, color=analogous)) +
  geom_density() +
  scale_color_brewer(palette="Set1") +
  facet_grid(~mode_rank)
```

```{r}
ggplot(summarized_sv_d, aes(x=epoch, y=mean_norm_S, color=analogous)) +
  geom_line() +
  scale_color_brewer(palette="Set1")
```

```{r}
ggplot(summarized_sv_d, aes(x=epoch, y=mean_S, color=analogous)) +
  geom_line() +
  scale_color_brewer(palette="Set1") +
  facet_grid(~mode_rank)
```

## do these features help explain transfer?
```{r}
learned_and_features_d = learned_d %>%
  filter(nonlinear, num_layers==3) %>%
  left_join(early_sv_d %>%
              filter(mode_rank==0)) %>%
  left_join(final_simil_d %>% 
              select(-mean_abs_cosine) %>%
              mutate(type=paste("cosine_distance_", type, sep="")) %>%
              spread(type, mean_cosine_distance)) %>%
  left_join(final_proj_d %>%
              filter(mode_rank==0))
```

```{r}
model = lmer(learned_epoch ~ analogous + S + cosine_distance_between_analog + cosine_distance_between_nonanalog + abs_proj_diff + (1 | run), learned_and_features_d)
BIC(model)
summary(model)
```

```{r}
simple_model = lmer(learned_epoch ~ analogous + S + (1 | run), learned_and_features_d)
BIC(simple_model)
summary(simple_model)
```

Seems like the early increase in S is doing much of the work, but not all (based on BIC difference).